Synthesis of Solar Photovoltaic Systems: Optimal Sizing Comparison\thanks{Supported by Newton Fund (ref. 261881580) and FAPEAM (Amazonas State Foundation for Research Support, calls 009/2017 and PROTI Pesquisa 2018).}

### 2020-04-23 Tex/CAV2019/Trindade_Cordeiro_VSTTE2020.tex
revised paper
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

\begin{abstract}
The use of power generation with renewable energy sources is proliferating due to industrial development. Photovoltaic (PV) panels have emerged as an alternative to fossil or nuclear fuel energy generation. The use of formal methods for PV systems is a new subject with significant research spanning only the last five years. Here we develop and evaluate a sound, automated synthesis approach to obtain optimal sizing PV systems. More specifically, we use a stand-alone configuration for this work since it is currently the most common system used for remote rural areas of developing countries or areas where the grid extension is not feasible. We propose a variant of the counterexample guided inductive synthesis (CEGIS) approach with two phases linking the technical and cost analysis. We advocate that the application of CEGIS to obtain optimal PV sizing has various advantages if compared to off-the-shelf optimization tools available in the market. Experimental results from seven case studies demonstrate that we can produce an optimal solution within an acceptable run-time; different verification tools are evaluated, in order to evaluate performance and soundness. We also compare our  CEGIS approach with a commercial tool specialized in PV systems optimization, with both results being validated with a complete and accurate commercial design software; furthermore, some real PV systems comparison are used in order to show our approach effectiveness. 
%First, we synthesize a feasible candidate based on power reliability, which may not attain the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. 
%However, the use of automated synthesis to obtain optimal sizing of solar photovoltaic systems remains unexplored in literature. 
\keywords{Formal symthesis \and Automated verification  \and Model Checking \and Solar photovoltaic systems.}
\end{abstract}

@@

Algorithm~\ref{alg:opt-algorithm} describes our pseudo-code to synthesize stand-alone PV systems using symbolic model checking~\cite{DBLP:journals/corr/abs-1909-13139}. The analytical method of optimization was adopted, with LCC economic analysis and power reliability based on critical period criteria (worst month's method).

 \begin{algorithm}
 \caption{Synthesis algorithm}
 \begin{algorithmic}[1]
 \renewcommand{\algorithmicrequire}{\textbf{Input:}}
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
  \REQUIRE weather data (temperature, solar irradiance); list of PV panels, controllers, batteries, and inverters datasheet information and cost; design requirements (load curve, peak demand, load surge power, daily energy consumption, battery autonomy, AC voltage); design assumptions (battery state of charge, depth of discharge, criteria and objectives for technical and cost analysis)
 \ENSURE FAIL (SAT) with counterexample showing the optimal sizing; SUCCESS (UNSAT), saying that the project has no feasible solution considering the requirements and the list of equipment
  \STATE Initialize variables \\
  \STATE Declare the maximum possible cost $MaxCost$  \\
  \STATE Calculate min possible Cost $MinCost$, based on the equipment list \\
  \FOR {$HintCost=MinCost$ to $MaxCost$}
 	\STATE Declare non-deterministic variable to select PV Panel from list \\
 	\STATE Declare non-deterministic variable to select Controller from list \\
 	\STATE Declare non-deterministic variable to select Battery from list \\
 	\STATE Declare non-deterministic variable to select Inverter from list \\ 	
 	\STATE Calculate $E_{corrected}, \, P_{min,panels} $ \\
	\STATE Calculate and define PV panels arrangement $N_{TP}, \, N_{PS}, N_{PP} $ \\
	\STATE Requirement enforced by \textbf{assume}$(P_{min,panels} \leq (N_{PS} \times N_{PP} \times P_{m,ref}))$ \\
	\STATE Calculate $I_{c,min} = N_{PP} \times I_{m,ref}$ \\
	\STATE Calculate $V_{c,min} = N_{PS} \times V_{m,ref}$ \\
 	\STATE Calculate $C_{bank}$ and $E_{b}$ \\
	\STATE Calculate and define battery arrangement $N_{BS}, \, N_{BP}, \, N_{BTotal}$ \\
	\STATE Requirement enforced by \textbf{assume} $(I_{min,DCbus} \leq (N_{BP} \times Capacity))$ \\
 	\STATE Controller requirements enforced by \textbf{assume}$((P_{controller} \geq P_{inverter}) \wedge V_{c} \wedge I_{c})$ \\
	\STATE Inverter requirements enforced by \textbf{assume}$(V_{in}DC \wedge V_{out}AC \wedge Demand \wedge P_{surge})$ \\
	\STATE non-deterministic variables hold feasible equipment and cost  \\
	\STATE $F_{obj} \leftarrow  N_{TP}*Panel_{Cost} \, + \, N_{TB}*Battery_{Cost} \, + Controller_{Cost} \, + \, Inverter_{Cost} \, + \, Installation_{Cost} \, + \, batrep_{Cost} \, + \, PWO\&M_{Cost}$ \\
	\STATE Violation check with \textbf{assert}$(F_{obj} > HintCost)$ \\
  \ENDFOR
 \RETURN $(\,)$ 
 \end{algorithmic} 
 \label{alg:opt-algorithm}
 \end{algorithm}

Our synthesis algorithm will synthesize constant values; 
it starts with the input of the manufacturer's data and prices of PV panels, batteries, charge controllers and inverters. Moreover, we define design (house) requirements and design assumptions.

@@

If our algorithm does not find a feasible solution among the item of equipment that were provided for our {\sc Synthesize} phase,  then the result is unsatisfiable (\textit{UNSAT}), i.e. the program finishes without finding a solution, indicating that it was not possible to combine the specific items of equipment in order to create a feasible solution.

The main challenge for the {\sc Synthesize} phase is to find a feasible candidate solution for the constraints and user requirements. The challenge for the {\sc Verify} phase is to find the lowest acquisition cost from a list of equipment and components provided by the {\sc Synthesize} phase.

Note that the process described here is completely automated and that a validation is performed by our {\sc Verify} phase to ensure that the approach is sound.

@@

In order to validate and compare the optimal sizing solution produced by our approach and by HOMER Pro, we use a complete simulation tool, called PVsyst version $6.86$, with plenty of commercial equipment in its data bank. We have considered a comparison for an entire year's weather data of simulation to guarantee that the proposed sizing meets the electrification requirements. PVsyst is a PC software package developed by a Swiss company used globally for the study, sizing, simulation, and data analysis of solar PV systems. PVsyst contains: design, sizing, 3D shading scene, simulation, grid, and off-grid features. It uses comprehensive irradiation data from Meteonorm\footnote{https://meteonorm.com/en/}, and aging analysis~\cite{PVsyst2017}. However, it does not perform optimization. Drawbacks of PVsyst: it does not perform optimization; therefore, it needs the system sized in order to validate it. PVsyst does not have commercial inverter equipment and therefore does not consider surge power demand as the ones produced by air conditioners and refrigerators for a few seconds. PVsyst is a paid software with $30$-day test possibility and runs only in MS-Windows operational system.

@@

The results are presented in Table~\ref{tab1}.

\begin{table}
\caption{Case studies and results: optimization of stand-alone PV systems.}\label{tab1}
\begin{scriptsize}
\begin{tabular}{c|c|c|c|c}
\hline
\hline
Tools & \makecell{CBMC 5.11 \\(MiniSat 2.2.1)}& \makecell{ESBMC 6.0.0 \\(Boolector 3.0.1)}& \makecell{CPAchecker 1.8\\(MathSAT 5.5.3)}& HOMER Pro 3.13.1\\
\hline
\hline
Specification & Result & Result & Result & Result \\
\hline
\makecell{\textbf{Case Study 1}\\Peak:342W\\Surge:342W \\E:3,900Wh/day\\Autonomy:48h} & OM & \makecell{SAT (620 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 700W/48V\\LCC: US\$ 10,214.04} & \makecell{SAT (548 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 700W/1600W/48V\\LCC: US\$ 10,214.04} & \makecell{(Time: 0.33 min)\\2.53 kW of PV\\NBT:12$\times$83.4Ah (2S-6P)\\0.351kW inverter\\LCC: US\$ 7,808.04}\\
\hline
\makecell{\textbf{Case Study 2}\\Peak:814W\\Surge:980W\\E:4,880Wh/day\\Autonomy:48h} & OM & TO & TO & \makecell{(Time: 0.18 min)\\3.71 kW of PV\\NBT:20$\times$83.4Ah (2S-10P)\\0.817kW inverter\\LCC: US\$ 12,861.75} \\
\hline
\makecell{\textbf{Case Study 3}\\Peak:815W\\Surge:980W\\E:4,880Wh/day\\Autonomy:12h} & OM & \makecell{SAT (63 min) \\NTP:14$\times$150W (7P-2S)\\NBT:6$\times$105Ah (2S-3P)\\Controller 35A/145V\\Inverter 1,200W/48V\\LCC: US\$ 9,274.07} & TO & Not possible \\
\hline
\makecell{\textbf{Case Study 4}\\Peak:253W\\Surge:722W\\E:3,600Wh/day\\Autonomy:48h} & OM & \makecell{SAT (147 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 280W/48V\\LCC: US\$ 9,678.63} & \makecell{SAT (605 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 280W/48V\\LCC: US\$ 9,678.63} & \makecell{(Time: 0.23 min)\\2.42 kW of PV\\NBT:12$\times$83.4Ah (2S-6P)\\0.254kW inverter\\LCC: US\$ 7,677.95}\\
\hline
\makecell{\textbf{Case Study 5}\\Peak:263W\\Surge:732W\\E:2,500Wh/day\\Autonomy:48h} & OM & \makecell {SAT (36.70 min) \\NTP:4$\times$330W (2S-2P)\\NBT:14$\times$80Ah (2S-7P)\\Controller 35A/145V\\Inverter 280W/24V \\LCC: US\$ 8,900.70}  & \makecell {SAT (254.25 min) \\NTP:4$\times$330W (2S-2P)\\NBT:14$\times$80Ah (2S-7P)\\Controller 35A/145V\\Inverter 280W/24V \\LCC: US\$ 8,900.70} & \makecell{(Time: 0.18 min)\\1.59 kW of PV\\NBT:10$\times$83.4Ah (2S-5P)\\0.268kW inverter\\LCC: US\$ 6,175.57} \\
\hline
\makecell{\textbf{Case Study 6}\\Peak:322W\\Surge:896W\\E:4,300Wh/day\\Autonomy:48h} & OM & \makecell {SAT (380.93 min) \\NTP:6$\times$320W (2P-3S)\\NBT:18$\times$105Ah (2S-9P)\\Controller 35A/145V\\Inverter 400W/24V \\LCC: US\$ 10,136.61} & TO & \makecell{(Time: 0.22 min)\\3.15 kW of PV\\NBT:14$\times$83.4Ah (2S-7P)\\0.328kW inverter\\LCC: US\$ 9,112.45} \\
\hline
\makecell{\textbf{Case Study 7}\\Peak:1,586W\\Surge:2,900W\\E:14,000Wh/day\\Autonomy:48h} & OM & UNSAT (0.48 min) & TO & \makecell{(Time: 0.20 min)\\12.5 kW of PV\\NBT:66$\times$83.4Ah (2S-33P)\\1.60kW inverter\\LCC: US\$ 41,878.11} \\
\hline
\hline
\end{tabular}
\\Legend: OM = out of memory; TO = time out; IF = internal failure; E = energy;        NTP = total number of panels, NBtotal = total number of batteries, NPS = number of panels in series; NPP = number of panels in parallel, NBS = number of batteries in series; NBP = number of batteries in parallel; LCC = Life Cycle Cost.
\end{scriptsize}
\end{table}

The violation (SAT result) indicated all over the Table~\ref{tab1} 
is the $assert$ of the line $21$ from Algorithm~\ref{alg:opt-algorithm} and indicates that an optimal solution was found.

CBMC was unable to produce any conclusive result. \textit{Out of memory} situations occurred in all case studies.

CPAchecker was able to synthesize optimal sizing in three out of the seven case studies (cases $1$, $4$, and $5$): the result was produced within the time limit, which varied from $254.25$ to $548$ minutes. Fig.~\ref{fig:CPAoptc1} illustrates the result of case $5$ with the optimal sizing appearing on the left side as the integer $3$ for the solar panel (which is the Canadian CS6U-330P model of $330$ W from the manufacturer Victron Energy), battery $0$ refers to the model 12MF80 of $80$ Ah from Moura, charge controller $0$ refers to the model 35A-145V MPPT from Victron Energy. The inverter number $2$ refers to the Epever model IP350-11 of $280$ W (nominal power) and $750$ W of surge power. The variables NTP, NPS, NPP, NBS, NBP, and NBTotal, also presented in the counterexample, shows the number of panels and batteries and how they are connected.

Case studies $2$, $3$, $6$, and $7$ led to a \textit{time out} result in CPAchecker, i.e. it was not solved within $660$ minutes.  
%
\begin{figure}[h]
\fbox{\includegraphics[width=1.0\textwidth]{CPA_opt_c5.png}}
\centering
\caption{Counterexample generated by CPAchecker after validation of case 5 (file Counterexample.html).}
\label{fig:CPAoptc1}
\end{figure}

An internal failure was presented by ESBMC when using Z3 solver in all case studies. It corresponds to a 'bug' in the Z3 solver, which will require an updated version of ESBMC to fix it. However, when using the incremental BMC configuration of ESBMC with Boolector solver, the verifier engine was able to reach the optimal sizing of case studies $1$, $3$, $4$, $5$, and $6$ with a FAIL/ SAT response varying from $36.70$ to $620$ minutes. Fig.~\ref{fig:esbmcoptcase3} shows the counterexample generated. ESBMC with this configuration was not possible to obtain an optimal solution in cases $2$ and $7$. Case $2$ reached a \textit{time out} result after $660$ minutes of processing. Moreover, case $7$ resulted in a UNSAT result, i.e., the verifier engine was not possible to get a feasible solution considering the list of equipment. This SAT situation was reached after just $0.48$ minutes.

\begin{figure}[h]
\fbox{\includegraphics[width=1.0\textwidth]{ESBMC_SAT_case3.png}}
\centering
\caption{ESBMC counterexample for case 3 optimization.}
\label{fig:esbmcoptcase3}
\end{figure}

Those specific results partially answers the \textit{EG2}.

HOMER Pro was able to evaluate six case studies (cases $1$, $2$, $4$, $5$, $6$, and $7$), and in under $30$ seconds, much faster than the proposed automated synthesis tool (cf.~\textit{EG3}). Case study $3$ could not be simulated since HOMER Pro does not have the battery autonomy adjustment feature, i.e., the tool always tries to feed the given load with electricity $365$ days/year.

Certain other HOMER Pro drawbacks were also noted: (1) System equipment does not include an explicit charge controller. HOMER Pro includes a controller automatically to simulate the charge/discharge of batteries and to meet the load requirement. However, without costs or even with electrical characteristics such as maximum current and voltage, which are common during PV sizing. (2) HOMER Pro requires the inclusion of some battery specification to initiate optimization; however, it does not change the electrical specifications during simulation; the results presented are multiples of the original battery type suggested by the user. For example, it was started with a $83.4$ Ah lead-acid battery, and during simulation, HOMER Pro did not try to use other capacities or types. (3) HOMER Pro does not present the optimal solution in terms of connections of PV panel arrays, just the total in terms of power, i.e., it presents neither the models and the power of each PV panel nor the total of panels in series or parallel.

@@

Comparing the results of the formal synthesis with CPAchecker, and ESBMC, against those of HOMER Pro, it was observed some distinct results, in terms of technical solution and cost (cf. Table~\ref{tab1}). Concerning the performance, there exists a vast difference in favor of HOMER Pro that obtained the results in considerably less time: few seconds in the opposite of an average of $234$ minutes for the automated synthesis technique.

Particularly in the case of LCC, the cost varied from $11$\% to $44$\%, always having a higher estimation from the automated synthesis technique. However, considering that the cost of individual items of each database used to compose the optimal design is not the same among the tools, it is plausible to obtain distinct results.

On the one hand, related to the PV panels sizing, the results presented by the automated synthesis were smaller in terms of power than the produced by the simulation tool. The difference varied from $19$\% to $65$\%. On the other hand, related to the battery bank, the results were smaller in terms of capacity to the HOMER Pro. The difference was between $34$\% to $68$\%. The mathematical models are different and particular parameters can be tuned as well in each technique, and that can justify the difference, which was presented in all the case studies.

Those discrepancies are not easy to address without some real systems validation. However, we use the simulation software PVsyst in order to validate the optimal sizing produced. According to PVsyst, we have to each case study:

>>>>

\begin{abstract}
The use of power generation with renewable energy sources is proliferating due to industrial development. Photovoltaic (PV) panels have emerged as an alternative to fossil or nuclear fuel energy generation. The use of formal methods for PV systems is a new subject with significant research spanning only the last five years. Here we develop and evaluate a sound, automated synthesis approach to obtain optimal sizing PV systems. More specifically, we use a stand-alone configuration for this work since it is currently the most common system used for remote rural areas of developing countries or areas where the grid extension is not feasible. We propose a variant of the counterexample guided inductive synthesis (CEGIS) approach with two phases linking the technical and cost analysis. We advocate that the application of CEGIS to obtain optimal PV sizing has various advantages if compared to off-the-shelf optimization tools available in the market. Experimental results from seven case studies demonstrate that we can produce an optimal solution within an acceptable run-time; different verification tools are evaluated, in order to evaluate performance and soundness. We also compare our  CEGIS approach with a commercial tool specialized in PV systems optimization, with both results being validated with a complete and accurate commercial design software; furthermore, some real PV systems comparison are used in order to show our approach effectiveness. 
%First, we synthesize a feasible candidate based on power reliability, which may not attain the lowest cost. Second, the candidate is then verified iteratively with a lower bound cost via model checking. If the verification step succeeds, the lower bound is adjusted; if it fails, a counterexample provides the optimal solution. 
%However, the use of automated synthesis to obtain optimal sizing of solar photovoltaic systems remains unexplored in literature. 
%\keywords{Formal symthesis \and Automated verification  \and Model Checking \and Solar photovoltaic systems.}
\end{abstract}

@@

Algorithm~\ref{alg:opt-algorithm} describes our pseudo-code to synthesize stand-alone PV systems using symbolic model checking~\cite{DBLP:journals/corr/abs-1909-13139}. The analytical method of optimization was adopted, with LCC economic analysis and power reliability based on critical period criteria (worst month's method).
%
 \begin{algorithm}
 \caption{Synthesis algorithm}
 \begin{algorithmic}[1]
 \renewcommand{\algorithmicrequire}{\textbf{Input:}}
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
  \REQUIRE weather data (temperature, solar irradiance); list of PV panels, controllers, batteries, and inverters datasheet information and cost; design requirements (load curve, peak demand, load surge power, daily energy consumption, battery autonomy, AC voltage); design assumptions (battery state of charge, depth of discharge, criteria and objectives for technical and cost analysis)
 \ENSURE FAIL (SAT) with counterexample showing the optimal sizing; SUCCESS (UNSAT), saying that the project has no feasible solution considering the requirements and the list of equipment
  \STATE Initialize variables \\
  \STATE Declare the maximum possible cost $MaxCost$  \\
  \STATE Calculate min possible Cost $MinCost$, based on the equipment list \\
  \FOR {$HintCost=MinCost$ to $MaxCost$}
 	\STATE Declare non-deterministic variable to select PV Panel from list \\
 	\STATE Declare non-deterministic variable to select Controller from list \\
 	\STATE Declare non-deterministic variable to select Battery from list \\
 	\STATE Declare non-deterministic variable to select Inverter from list \\ 	
 	\STATE Calculate $E_{corrected}, \, P_{min,panels} $ \\
	\STATE Calculate and define PV panels arrangement $N_{TP}, \, N_{PS}, N_{PP} $ \\
	\STATE Requirement enforced by \textbf{assume}$(P_{min,panels} \leq (N_{PS} \times N_{PP} \times P_{m,ref}))$ \\
	\STATE Calculate $I_{c,min} = N_{PP} \times I_{m,ref}$ \\
	\STATE Calculate $V_{c,min} = N_{PS} \times V_{m,ref}$ \\
 	\STATE Calculate $C_{bank}$ and $E_{b}$ \\
	\STATE Calculate and define battery arrangement $N_{BS}, \, N_{BP}, \, N_{BTotal}$ \\
	\STATE Requirement enforced by \textbf{assume} $(I_{min,DCbus} \leq (N_{BP} \times Capacity))$ \\
 	\STATE Controller requirements enforced by \textbf{assume}$((P_{controller} \geq P_{inverter}) \wedge V_{c} \wedge I_{c})$ \\
	\STATE Inverter requirements enforced by \textbf{assume}$(V_{in}DC \wedge V_{out}AC \wedge Demand \wedge P_{surge})$ \\
	\STATE non-deterministic variables hold feasible equipment and cost  \\
	\STATE $F_{obj} \leftarrow  N_{TP}*Panel_{Cost} \, + \, N_{TB}*Battery_{Cost} \, + Controller_{Cost} \, + \, Inverter_{Cost} \, + \, Installation_{Cost} \, + \, batrep_{Cost} \, + \, PWO\&M_{Cost}$ \\
	\STATE Violation check with \textbf{assert}$(F_{obj} > HintCost)$ \\
  \ENDFOR
 \RETURN $(\,)$ 
 \end{algorithmic} 
 \label{alg:opt-algorithm}
 \end{algorithm}
%
Our synthesis algorithm will synthesize constant values; 
it starts with the input of the manufacturer's data and prices of PV panels, batteries, charge controllers and inverters. Moreover, we define design (house) requirements and design assumptions.

@@

If our algorithm does not find a feasible solution among the item of equipment that were provided for our {\sc Synthesize} phase,  then the result is unsatisfiable (\textit{UNSAT}), i.e., the program finishes without finding a solution, indicating that it was unable to combine the specific items of equipment to create a feasible solution. 
%
The main challenge for the {\sc Synthesize} phase is to find a feasible candidate solution for the constraints and user requirements. The challenge for the {\sc Verify} phase is to find the lowest acquisition cost from a list of equipment and components provided by the {\sc Synthesize} phase. 
%
Note that the process described here is completely automated and that a validation is performed by our {\sc Verify} phase to ensure that the approach is sound.

@@

In order to validate and compare the optimal sizing solution produced by our approach and by HOMER Pro, we use a simulation tool, called PVsyst version $6.86$, with plenty of commercial equipment in its data bank. We have considered a comparison for an entire year's weather data of simulation to guarantee that the proposed sizing meets the electrification requirements. PVsyst is a PC software package developed by a Swiss company used for the study, sizing, simulation, and data analysis of solar PV systems. PVsyst contains: design, sizing, 3D shading scene, simulation, grid, and off-grid features. It uses comprehensive irradiation data from Meteonorm,\footnote{https://meteonorm.com/en/}, and aging analysis~\cite{PVsyst2017}. However, it does not perform optimization; therefore, PVsyst needs the system sized in order to validate it. Furthermore, PVsyst does not have commercial inverter equipment and, as a result, does not consider surge power demand as the ones produced by air conditioners and refrigerators for a few seconds. PVsyst is a paid software with $30$-day test possibility and runs only in MS-Windows operational system.

@@

The results are presented in Table~\ref{tab1}. 
%
\begin{table}
\caption{Case studies and results: optimization of stand-alone PV systems.}\label{tab1}
\begin{scriptsize}
\begin{tabular}{c|c|c|c|c}
\hline
\hline
Tools & \makecell{CBMC 5.11 \\(MiniSat 2.2.1)}& \makecell{ESBMC 6.0.0 \\(Boolector 3.0.1)}& \makecell{CPAchecker 1.8\\(MathSAT 5.5.3)}& HOMER Pro 3.13.1\\
\hline
\hline
Specification & Result & Result & Result & Result \\
\hline
\makecell{\textbf{Case Study 1}\\Peak:342W\\Surge:342W \\E:3,900Wh/day\\Autonomy:48h} & OM & \makecell{SAT (620 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 700W/48V\\LCC: US\$ 10,214.04} & \makecell{SAT (548 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 700W/1600W/48V\\LCC: US\$ 10,214.04} & \makecell{(Time: 0.33 min)\\2.53 kW of PV\\NBT:12$\times$83.4Ah (2S-6P)\\0.351kW inverter\\LCC: US\$ 7,808.04}\\
\hline
\makecell{\textbf{Case Study 2}\\Peak:814W\\Surge:980W\\E:4,880Wh/day\\Autonomy:48h} & OM & TO & TO & \makecell{(Time: 0.18 min)\\3.71 kW of PV\\NBT:20$\times$83.4Ah (2S-10P)\\0.817kW inverter\\LCC: US\$ 12,861.75} \\
\hline
\makecell{\textbf{Case Study 3}\\Peak:815W\\Surge:980W\\E:4,880Wh/day\\Autonomy:12h} & OM & \makecell{SAT (63 min) \\NTP:14$\times$150W (7P-2S)\\NBT:6$\times$105Ah (2S-3P)\\Controller 35A/145V\\Inverter 1,200W/48V\\LCC: US\$ 9,274.07} & TO & Not possible \\
\hline
\makecell{\textbf{Case Study 4}\\Peak:253W\\Surge:722W\\E:3,600Wh/day\\Autonomy:48h} & OM & \makecell{SAT (147 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 280W/48V\\LCC: US\$ 9,678.63} & \makecell{SAT (605 min) \\NTP:6$\times$330W (2P-3S)\\NBT:16$\times$105Ah (2S-8P)\\Controller 35A/145V\\Inverter 280W/48V\\LCC: US\$ 9,678.63} & \makecell{(Time: 0.23 min)\\2.42 kW of PV\\NBT:12$\times$83.4Ah (2S-6P)\\0.254kW inverter\\LCC: US\$ 7,677.95}\\
\hline
\makecell{\textbf{Case Study 5}\\Peak:263W\\Surge:732W\\E:2,500Wh/day\\Autonomy:48h} & OM & \makecell {SAT (36.70 min) \\NTP:4$\times$330W (2S-2P)\\NBT:14$\times$80Ah (2S-7P)\\Controller 35A/145V\\Inverter 280W/24V \\LCC: US\$ 8,900.70}  & \makecell {SAT (254.25 min) \\NTP:4$\times$330W (2S-2P)\\NBT:14$\times$80Ah (2S-7P)\\Controller 35A/145V\\Inverter 280W/24V \\LCC: US\$ 8,900.70} & \makecell{(Time: 0.18 min)\\1.59 kW of PV\\NBT:10$\times$83.4Ah (2S-5P)\\0.268kW inverter\\LCC: US\$ 6,175.57} \\
\hline
\makecell{\textbf{Case Study 6}\\Peak:322W\\Surge:896W\\E:4,300Wh/day\\Autonomy:48h} & OM & \makecell {SAT (380.93 min) \\NTP:6$\times$320W (2P-3S)\\NBT:18$\times$105Ah (2S-9P)\\Controller 35A/145V\\Inverter 400W/24V \\LCC: US\$ 10,136.61} & TO & \makecell{(Time: 0.22 min)\\3.15 kW of PV\\NBT:14$\times$83.4Ah (2S-7P)\\0.328kW inverter\\LCC: US\$ 9,112.45} \\
\hline
\makecell{\textbf{Case Study 7}\\Peak:1,586W\\Surge:2,900W\\E:14,000Wh/day\\Autonomy:48h} & OM & UNSAT (0.48 min) & TO & \makecell{(Time: 0.20 min)\\12.5 kW of PV\\NBT:66$\times$83.4Ah (2S-33P)\\1.60kW inverter\\LCC: US\$ 41,878.11} \\
\hline
\hline
\end{tabular}
\\Legend: OM = out of memory; TO = time out; IF = internal failure; E = energy;        NTP = total number of panels, NBtotal = total number of batteries, NPS = number of panels in series; NPP = number of panels in parallel, NBS = number of batteries in series; NBP = number of batteries in parallel; LCC = Life Cycle Cost.
\end{scriptsize}
\end{table}
%
The violation (SAT result) indicated all over the Table~\ref{tab1} 
is the $assert$ of the line $21$ from Algorithm~\ref{alg:opt-algorithm} and indicates that an optimal solution was found.
Here, CBMC was unable to produce any conclusive result. \textit{Out of memory} situations occurred in all case studies.

CPAchecker was able to synthesize optimal sizing in three out of the seven case studies (cases $1$, $4$, and $5$): the result was produced within the time limit, which varied from $4$ to $9$ hours. 
Fig.~\ref{fig:CPAoptc1} illustrates the result of case $5$ with the optimal sizing appearing on the left side as the integer $3$ for the solar panel (which is the Canadian CS6U-330P model of $330$ W from the manufacturer Victron Energy), battery $0$ refers to the model 12MF80 of $80$ Ah from Moura, charge controller $0$ refers to the model 35A-145V MPPT from Victron Energy. The inverter number $2$ refers to the Epever model IP350-11 of $280$ W (nominal power) and $750$ W of surge power. The variables NTP, NPS, NPP, NBS, NBP, and NBTotal, also presented in the counterexample, shows the number of panels and batteries and how they are connected.
Case studies $2$, $3$, $6$, and $7$ led to a \textit{time out} result in CPAchecker, i.e., it was not solved within $660$ minutes.  
%
\begin{figure}[h]
\fbox{\includegraphics[width=1.0\textwidth]{CPA_opt_c5.png}}
\centering
\caption{Counterexample generated by CPAchecker after validation of case $5$.}
\label{fig:CPAoptc1}
\end{figure}

%An internal failure was presented by ESBMC when using Z3 solver in all case studies. It corresponds to a 'bug' in the Z3 solver, which will require an updated version of ESBMC to fix it.

Here we used ESBMC in incremental BMC configuration with Boolector; ESBMC was able to reach the optimal sizing of case studies $1$, $3$, $4$, $5$, and $6$ with a FAIL/ SAT response, varying from $36$ minutes to $10$ hours. 
%Fig.~\ref{fig:esbmcoptcase3} shows the counterexample generated. 
ESBMC with this configuration was unable to obtain an optimal solution in cases $2$ and $7$. Case $2$ reached a \textit{time out} result after $11$ hours. Moreover, case $7$ resulted in a UNSAT result, i.e., the verifier engine was unable to provide a feasible solution considering the equipment list \textcolor{red}{why? is it a bug in ESBMC?}. This UNSAT situation was reached in less than one minute.
%
%\begin{figure}[h]
%\fbox{\includegraphics[width=1.0\textwidth]{ESBMC_SAT_case3.png}}
%\centering
%\caption{ESBMC counterexample for case 3 optimization.}
%\label{fig:esbmcoptcase3}
%\end{figure}
%
These experimental results partially answers the \textit{EG2} \textcolor{red}{why ``partially''? explain here}.

HOMER Pro was able to evaluate six case studies (cases $1$, $2$, $4$, $5$, $6$, and $7$) under $30$ seconds; it was much faster than the proposed automated synthesis tool (cf.~\textit{EG3}). Case study $3$ could not be simulated since HOMER Pro does not have the battery autonomy adjustment feature, i.e., the tool always tries to feed the given load with electricity $365$ days/year.

Some HOMER Pro drawbacks were also noted: (1) System equipment does not include an explicit charge controller. HOMER Pro includes a controller automatically to simulate the charge/discharge of batteries and to meet the load requirement. However, without costs or even with electrical characteristics such as maximum current and voltage, which are common during PV sizing. (2) HOMER Pro requires the inclusion of some battery specification to initiate optimization; however, it does not change the electrical specifications during simulation; the results presented are multiples of the original battery type suggested by the user. For example, it was started with a $83.4$ Ah lead-acid battery, and during simulation, HOMER Pro did not try to use other capacities or types. (3) HOMER Pro does not present the optimal solution in terms of connections of PV panel arrays, just the total in terms of power, i.e., it presents neither the models and the power of each PV panel nor the total of panels in series or parallel.

@@

If we compare the results of the formal synthesis with CPAchecker and ESBMC against those of HOMER Pro, we observed some distinct results in terms of the technical solution and cost (cf. Table~\ref{tab1}). Concerning the performance, there exists a vast difference in favor of HOMER Pro that obtained the results in considerably less time: few seconds in the opposite of an average of $4$ hours for the automated synthesis technique.

Particularly in the case of LCC, the cost varied from $11$\% to $44$\%, always producing a higher estimation from the automated synthesis technique. However, considering that the cost of individual items of each database used to compose the optimal design is not the same among the tools, it is plausible to obtain distinct results.

On the one hand, concerning the PV panels sizing, the results presented by the automated synthesis were smaller in terms of power than the ones produced by the simulation tool. The difference varied from $19$\% to $65$\%. On the other hand, concerning the battery bank, the results were smaller in terms of capacity to the HOMER Pro. The difference was between $34$\% to $68$\%. The mathematical models are different and particular parameters can be tuned as well on each technique, and that can justify the difference, which was presented in all the case studies.

Those discrepancies are not easy to address without some real systems validation. However, we use the simulation software PVsyst to validate the optimal sizing produced. According to PVsyst, we have to each case study:
